Ring of Fire Camera Detections Report

### 2023-02-28 analysis/paper/paper.qmd
updates

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  dplyr::filter(!is.na(latitude)) %>% 
  dplyr::filter(!is.na(longitude)) %>% 
  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
  dplyr::mutate(is_retrieved = is.na(date_retrieved))

@@

```{r}
#| label: ordination
#| echo: false

@@

```{r}
#| label: ordination
#| echo: false

@@

```{r}
#| label: ordination
#| echo: false

@@

```{r}
events_tb_cam <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(camera_id, species, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct() %>% 
  dplyr::group_by(camera_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup() %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(camera_id) %>%
  tibble::column_to_rownames("camera_id")

@@

```{r}
clust <- hclust(cluster::daisy(events_tb_cam_hell, "gower"), 
                method = "ward.D2")
plot(clust)
```

```{r}
clust_grp <- cutree(clust, k=5)
custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_grp)
```

```{r}
coords <- cams_sf %>% 
  sf::st_coordinates() %>% 
  as.data.frame()
cams_sf_with_props <- dplyr::bind_cols(cams_sf, coords) %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(plot_id,X,Y) %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
  dplyr::distinct() %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(events, by = "plot_id")

@@

```{r}
zones <- 
  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp") %>% 
  sf::st_transform(crs = sf::st_crs(cams_sf))

@@

```{r}
zones_filter <- zones %>%
  sf::st_filter(cams_sf)
plot(zones_filter)
```

```{r}
ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf()
```

```{r}
ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
                   ggplot2::aes(col = is_retrieved)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

```{r}
pols <- sf::st_coordinates(zones_filter) %>% 
  as.data.frame()

@@

for (pol in 1:3){
  base <- base + 
    ggplot2::geom_path(data = dplyr::filter(pols, L3 == pol),
                       ggplot2::aes(x=X, y=Y),
                       linewidth = 0.2)
}

base +
  ggplot2::coord_equal(xlim = c(-88, -84.5),
                       ylim = c(50, 54)) +
  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION), 
                      data = regions_labels) +
  scatterpie::geom_scatterpie(data = cams_sf_with_props,
                              ggplot2::aes(x=X, y=Y, group=plot_id),
                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
                              sorted_by_radius = TRUE,
                              legend_name = "Species",
                              pie_scale = 1.5)
```

```{r}
set.seed(123)
long <- rnorm(50, sd=100)
lat <- rnorm(50, sd=50)
d <- data.frame(long=long, lat=lat)
d <- with(d, d[abs(long) < 150 & abs(lat) < 70,])
n <- nrow(d)
d$region <- factor(1:n)
d$A <- abs(rnorm(n, sd=1))
d$B <- abs(rnorm(n, sd=2))
d$C <- abs(rnorm(n, sd=3))
d$D <- abs(rnorm(n, sd=4))
d[1, 4:7] <- d[1, 4:7] * 3
head(d)
ggplot2::ggplot() + scatterpie::geom_scatterpie(ggplot2::aes(x=long, y=lat, group=region), data=d,
                           cols=LETTERS[1:4]) + ggplot2::coord_equal()
```

>>>>

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  dplyr::filter(!is.na(latitude)) %>% 
  dplyr::filter(!is.na(longitude)) %>% 
  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
  dplyr::mutate(is_retrieved = !is.na(date_retrieved))

@@

```{r}
#| label: ordination-base
#| echo: false

@@

```{r}
#| label: ordination-custom
#| echo: false

@@

```{r}
#| label: events
#| echo: false

@@

```{r}
#| label: ordination-events
#| echo: false

anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

events_tb_cam <- anns_wide_animals_joined_mut %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(camera_id, species, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct() %>% 
  dplyr::group_by(camera_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup() %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(camera_id) %>%
  tibble::column_to_rownames("camera_id")

@@

```{r}
#| label: cluster-base
#| echo: false

events_gower <- cluster::daisy(events_tb_cam_hell, "gower")
clust <- hclust(events_gower, method = "ward.D2")
plot(clust)
```

```{r}
#| label: rda-clust
#| echo: false

clust_grp <- cutree(clust, k=5)
custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_grp) 
```

```{r}
#| label: test-pie
#| echo: false

coords <- cams_sf %>% 
  sf::st_coordinates() %>% 
  as.data.frame()

cams_sf_coords <- dplyr::bind_cols(cams_sf, coords)

cams_sf_with_props <- cams_sf_coords %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(plot_id,X,Y) %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
  dplyr::distinct() %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(events, by = "plot_id")

@@

```{r}
#| label: ecoregions
#| echo: false

zones <- 
  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp") %>% 
  sf::st_transform(crs = sf::st_crs(cams_sf))

@@

```{r}
#| label: ecoregions-basic
#| echo: false

zones_filter <- zones %>%
  sf::st_filter(cams_sf)
plot(zones_filter)
```

```{r}
#| label: ecoregions-ggplot
#| echo: false

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf()
```

```{r}
#| label: ecoregions-ggplot-retrieved
#| echo: false

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
                   ggplot2::aes(col=is_retrieved)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

```{r}
#| label: map-pies
#| echo: false

pols <- sf::st_coordinates(zones_filter) %>% 
  as.data.frame()

@@

for (pol in 1:3){
  base <- base + 
    ggplot2::geom_polygon(data = dplyr::filter(pols, L3 == pol),
                          ggplot2::aes(x=X, y=Y),
                          col = 1, fill = "grey90")
}

base +
  ggplot2::coord_equal(xlim = c(-88, -84.5),
                       ylim = c(50, 54)) +
  ggplot2::geom_label(ggplot2::aes(x=x, y=y, label=ECOREGION), 
                      data = regions_labels) +
  scatterpie::geom_scatterpie(data = cams_sf_with_props,
                              ggplot2::aes(x=X, y=Y, group=plot_id),
                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)],
                              sorted_by_radius = TRUE,
                              legend_name = "Species",
                              pie_scale = 1.5) +
  ggplot2::labs(x="Lat", y="Long")
```

```{r}
#| label: map-groups
#| echo: false

grps_df <- data.frame(camera_id = names(clust_grp),
                      grp = as.factor(clust_grp))

cams_sf_clust <- cams_sf %>% 
  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
  dplyr::left_join(grps_df, by = "camera_id") %>% 
  dplyr::filter(!(is.na(grp)))

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf_clust, cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

```{r}
#| label: clust-geo
#| echo: false

# Spatially constrained clust

events_gower_mod <- events_gower
class(events_gower_mod) <- "dist"

tree <- ClustGeo::hclustgeo(events_gower_mod)
# plot(tree)
# rect.hclust(tree ,k = 5, border = c(4,5,3,2,1))
# legend("topright", legend = paste("cluster",1:5), 
#        fill=1:5,bty= "n", border = "white")

dist_geo <- cams_sf_clust %>% 
  dplyr::arrange(camera_id) %>% 
  sf::st_distance() %>% 
  as.dist()

range.alpha <- seq(0,1,0.1)
K <- 5
cr <- ClustGeo::choicealpha(events_gower_mod, dist_geo, range.alpha, 
  K, graph = F)

clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
                                 dist_geo, alpha = 1)
clust_geo_grp <- cutree(clust_geo, k=K)

grps_geo_df <- data.frame(camera_id = names(clust_geo_grp),
                          grp = as.factor(clust_geo_grp))

cams_sf_geo_clust <- cams_sf %>% 
  # dplyr::filter(camera_id %in% names(clust_grp)) %>% 
  dplyr::left_join(grps_geo_df, by = "camera_id") %>% 
  dplyr::filter(!(is.na(grp)))

plot(clust_geo)
```

```{r}
#| label: clust-geo-plot
#| echo: false

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf_geo_clust, cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

```{r}
#| label: rda-clust-geo
#| echo: false

custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
```

### Density heatmaps

```{r}
#| label: join-sf
#| echo: false

cams_sf_sp <- anns_wide_animals_joined_mut %>% 
  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
  dplyr::filter(species %in% colnames(events[,-1]))

cams_sf_sp_sum <- cams_sf_sp %>% 
  dplyr::group_by(camera_id, X, Y, species, month) %>% 
  dplyr::summarise(count=dplyr::n()) %>% 
  dplyr::ungroup()
# stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])

# cams_sf_sp %>% 
cams_sf_sp_sum %>% 
  # dplyr::filter(species %in% colnames(events[,-1])) %>% 
  dplyr::filter(species == "Caribou") %>% 
  # dplyr::filter(species == "Caribou", month == 8) %>% 
  ggplot2::ggplot() +
  # ggplot2::geom_hex(ggplot2::aes(x=X, y=Y), bins = 70) +
  # ggplot2::stat_density_2d(ggplot2::aes(fill = ..level..), geom = "polygon", colour="white") +
  # ggplot2::geom_density_2d(ggplot2::aes(x=X, y=Y)) +
  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
  #                          geom = "raster", contour = T) +
  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
  #                          geom = "raster", contour = F) +
  # ggplot2::geom_tile(ggplot2::aes(x=X, y=Y, fill = count),interpolate=TRUE) +
  ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size = count)) +
  # ggplot2::scale_fill_continuous(type = "viridis") +
  # ggplot2::scale_fill_distiller(palette=4, direction=1) +
  # ggplot2::theme_bw() + 
  ggplot2::facet_wrap(~month) #+
  # ggplot2::scale_x_continuous(expand = c(0, 0)) +
  # ggplot2::scale_y_continuous(expand = c(0, 0)) # +
  # ggplot2::theme(
  #   legend.position='none'
  # )
```

```{r}
ggplot2::ggplot(ggplot2::faithfuld, ggplot2::aes(waiting, eruptions)) +
 ggplot2::geom_raster(ggplot2::aes(fill = density))
```
